### 2016-06-24 analysis.Rmd
Fix catch error

---
title: "Psych 254 Replication Project Analysis"
author: "Eric Smith, Robert Hawkins, Michael C. Frank"
date: '`r Sys.Date()`'
output:
  html_document:
    number_sections: yes
    toc: yes
---

Libraries.

```{r}
library(dplyr)
library(ggplot2)
library(googlesheets)
library(readr)
library(tidyr)
```

Get project planning sheet straight from google docs.

```{r}
project_plan_sheet <- gs_url("https://docs.google.com/spreadsheets/d/1KZw75xlypvtK0cDoOHqXfFd3HsrfJ4gbq8iEJ6LOa0Y", lookup = TRUE) # Need lookup=T to initialize
project_plan_raw <- gs_read_csv(project_plan_sheet)
```

Exclude those to be excluded.

```{r}
project_plan <- project_plan_raw %>%
  filter(`include_in_final_writeup`==1) 
```

Read in project_info data from each project

```{r}
errors <- list()# Initialize an error var

all_project_info <- data.frame( # initialize an empty dataframe
  project_key = character()
)

project_keys <- project_plan$project_key # Find user names that are included in analysis
# For each user_

temp_project_keys <- c("smraposo","jreynolds","jmarias","ehermann", "mkeil") # Only do users we have data for currently

# Run all respective R code - need loop because project_keys same variable name
rm("project_info") # In case var is saved
for(project_key in project_keys) { # change from temp - test; project_key <- "smraposo"
  print(paste0("Running ",project_key))
  fp <- paste0("project_analyses/",project_key,".R")
  #source(fp)
  tryCatch(source(fp), error=function(e) {print(e)}) 
  if(exists("project_info")) { # If has project_info dataframe
    all_project_info <- bind_rows(all_project_info,project_info) # add to all data
    rm("project_info") # In case other file does not same
  } else {
    errors <- c(errors,paste0(project_key," is missing project_info"))
  }
}

```